---
title: "Import - Edge angle analysis - experimental data"
author: "Lisa Schunk"
date: "`r Sys.time()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: cerulean
    highlight: pygments
    number_sections: true
  github_document: 
    toc: true
    toc_depth: 3
    html_preview: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_format = "all", knit_root_dir = rprojroot::find_rstudio_root_file()) })

---

```{r Knitr Options, include = FALSE}
knitr::opts_chunk$set(comment = NA, message = FALSE, indent = "", error = TRUE)
```


---

# Goal of the script
This script imports and merges all single CSV-files generated with the 'edge angle method'. The data derives from 3D models of artefacts from three different experiments: Initial experiment, arificial VS natural experiment, tool function experiment. The data always cotains the 3D models from before, after 50, 250, 1000 and 2000 strokes.

The script will:

1. Read in the original CSV-files   
2. Combine the data from all samples into one 
3. Write an XLSX-file and save an R object ready for further analysis in R 



```{r}
dir_in <- "analysis/raw_data"
dir_out <- "analysis/derived_data/"
```


Raw data must be located in "`r dir_in`".  
Formatted data will be saved in "`r dir_out`".
The knit directory for this script is the project directory.

---


# Load packages
```{r}
library(tidyverse)
library(R.utils)
library(openxlsx)
library(tools)
```


---


# List all files and get names of the files  
```{r}
# List all CSV files in dir_in
CSV_files <- list.files(dir_in, pattern = "\\.csv$", recursive = TRUE, full.names = TRUE)

```



# Merge all files and format the data 
```{r}
# Create a list 
data_final <- vector(mode = "list", length = length(CSV_files)) 
names(data_final) <- basename(CSV_files)

# For each sample
for (s in seq_along(data_final)) {
  
      
    # Gets name of the experiment from path names 
    exp <- dirname(dirname(dirname(CSV_files[s]))) %>%
          basename()
    
   # read the data files
   data_final[[s]] <- read.csv(CSV_files[s]) %>% 
                      mutate(experiment = exp) %>% 
                      select(experiment, everything()) %>% 
                      rename(Angle_number = angel_number, Distance_origin = dist.to.origin.on.curve..mm., Segment = segment.on.section..mm.,Three_point = angle.1..3.points...degree., Two_lines = angle.2..2.constructed.lines...degree., Best_fit = angle.3..2.BestFit.lines...degree.)
}

# rbind all files 
data_final2 <- do.call(rbind, data_final)
# adds indices as row names 
row.names(data_final2) <- 1:nrow(data_final2)

# split column section
underscore_split <- strsplit(data_final2[["section"]], "_")
underscore_bind <- do.call(rbind, underscore_split)
minus_split <- strsplit(underscore_bind[,1], "-")
minus_bind <- do.call(rbind, minus_split)
sample_ID <- paste(minus_bind[,1], minus_bind[,2], sep="-")

data_final3 <- data_final2 %>% 
  mutate(ID = sample_ID, strokes = minus_bind[,3], edge = underscore_bind[,2], sec = underscore_bind[,4])

# extracts the raw material based on the ID
data_final3[grep("FLT", data_final3[["ID"]]), "Raw.material"] <- "flint"
data_final3[grep("LYDIT", data_final3[["ID"]]), "Raw.material"] <- "lydite"
data_final3[["Raw.material"]] <- factor(data_final3[["Raw.material"]])

data_final3[["Raw.material"]] <- factor(data_final3[["Raw.material"]])

# adds the contact/worked material
data_final3[grep("LYDIT4-1", data_final3[["ID"]]), "Contact.material"] <- "pork skin"
data_final3[grep("LYDIT4-2", data_final3[["ID"]]), "Contact.material"] <- "bone plate"
data_final3[grep("LYDIT4-5", data_final3[["ID"]]), "Contact.material"] <- "bos scapula"
data_final3[grep("LYDIT4-9", data_final3[["ID"]]), "Contact.material"] <- "skin pad"

data_final3[grep("FLT4-4", data_final3[["ID"]]), "Contact.material"] <- "pork skin"
data_final3[grep("FLT4-5", data_final3[["ID"]]), "Contact.material"] <- "bone plate"
data_final3[grep("FLT4-15", data_final3[["ID"]]), "Contact.material"] <- "bos scapula"
data_final3[grep("FLT4-12", data_final3[["ID"]]), "Contact.material"] <- "skin pad"

data_final3[grep("initial_experiment", data_final3[["experiment"]]), "Contact.material"] <- "bone plate"
data_final3[grep("tool_function-experiment_cutting", data_final3[["experiment"]]), "Contact.material"] <- "bone plate"
data_final3[grep("tool_function-experiment_carving", data_final3[["experiment"]]), "Contact.material"] <- "bone plate"

data_final3[["experiment"]] <- factor(data_final3[["experiment"]])

# adds column about the task/movement 
data_final3[grep("tool_function-experiment_cutting", data_final3[["experiment"]]), "Task"] <- "cutting"
data_final3[grep("tool_function-experiment_carving", data_final3[["experiment"]]), "Task"] <- "carving"
data_final3[grep("initial_experiment", data_final3[["experiment"]]), "Task"] <- "cutting"
data_final3[grep("'aVSn'-experiment", data_final3[["experiment"]]), "Task"] <- "cutting"
data_final3[grep("LYDIT5-14", data_final3[["ID"]]), "Task"] <- "scraping"

data_final3[["Task"]] <- factor(data_final3[["Task"]])

# adds column about the edge angle 
data_final3[grep("'aVSn'-experiment", data_final3[["experiment"]]), "Edge.angle"] <- "60°"

data_final3[grep("FLT8-1", data_final3[["ID"]]), "Edge.angle"] <- "45°"
data_final3[grep("FLT8-2", data_final3[["ID"]]), "Edge.angle"] <- "45°"
data_final3[grep("FLT8-3", data_final3[["ID"]]), "Edge.angle"] <- "45°"
data_final3[grep("LYDIT5-2", data_final3[["ID"]]), "Edge.angle"] <- "45°"
data_final3[grep("LYDIT5-3", data_final3[["ID"]]), "Edge.angle"] <- "45°"
data_final3[grep("LYDIT5-4", data_final3[["ID"]]), "Edge.angle"] <- "45°"
data_final3[grep("FLT8-4", data_final3[["ID"]]), "Edge.angle"] <- "35°"
data_final3[grep("FLT8-5", data_final3[["ID"]]), "Edge.angle"] <- "35°"
data_final3[grep("FLT8-6", data_final3[["ID"]]), "Edge.angle"] <- "35°"
data_final3[grep("LYDIT5-5", data_final3[["ID"]]), "Edge.angle"] <- "35°"
data_final3[grep("LYDIT5-6", data_final3[["ID"]]), "Edge.angle"] <- "35°"
data_final3[grep("LYDIT5-7", data_final3[["ID"]]), "Edge.angle"] <- "35°"

data_final3[grep("FLT8-10", data_final3[["ID"]]), "Edge.angle"] <- "45°"
data_final3[grep("FLT8-11", data_final3[["ID"]]), "Edge.angle"] <- "45°"
data_final3[grep("FLT8-12", data_final3[["ID"]]), "Edge.angle"] <- "45°"
data_final3[grep("LYDIT5-8", data_final3[["ID"]]), "Edge.angle"] <- "45°"
data_final3[grep("LYDIT5-9", data_final3[["ID"]]), "Edge.angle"] <- "45°"
data_final3[grep("LYDIT5-10", data_final3[["ID"]]), "Edge.angle"] <- "45°"
data_final3[grep("FLT8-7", data_final3[["ID"]]), "Edge.angle"] <- "35°"
data_final3[grep("FLT8-8", data_final3[["ID"]]), "Edge.angle"] <- "35°"
data_final3[grep("FLT8-9", data_final3[["ID"]]), "Edge.angle"] <- "35°"
data_final3[grep("LYDIT5-11", data_final3[["ID"]]), "Edge.angle"] <- "35°"
data_final3[grep("LYDIT5-12", data_final3[["ID"]]), "Edge.angle"] <- "35°"
data_final3[grep("LYDIT5-13", data_final3[["ID"]]), "Edge.angle"] <- "35°"
data_final3[grep("LYDIT5-14", data_final3[["ID"]]), "Edge.angle"] <- "35°"

data_final3[grep("FLT4-2", data_final3[["ID"]]), "Edge.angle"] <- "40°"
data_final3[grep("FLT4-3", data_final3[["ID"]]), "Edge.angle"] <- "60°"
data_final3[grep("LYDIT1-2", data_final3[["ID"]]), "Edge.angle"] <- "40°"
data_final3[grep("LYDIT1-3", data_final3[["ID"]]), "Edge.angle"] <- "40°"
data_final3[grep("LYDIT1-4", data_final3[["ID"]]), "Edge.angle"] <- "40°"
data_final3[grep("LYDIT3-2", data_final3[["ID"]]), "Edge.angle"] <- "60°"
data_final3[grep("LYDIT3-3", data_final3[["ID"]]), "Edge.angle"] <- "60°"

data_final3[["Edge.angle"]] <- factor(data_final3[["Edge.angle"]])


# reorder columns
data_final3 <- data_final3[c(2, 1, 9:16, 3:8)]

```



# Save data
## Format name of output file
```{r}
file_out <- "EdgeAngle_experiment"
```


## Write to XLSX
```{r}
write.xlsx(list(data = data_final3), file = paste0(dir_out, file_out, ".xlsx")) 
```


## Save R object
```{r}
saveObject(data_final, file = paste0(dir_out, file_out, ".Rbin"))
```


## Show files information
```{r}
files_out <- list.files(dir_out, full.names = TRUE)
md5_out <- md5sum(files_out)
info_out <- data.frame(files = basename(names(md5_out)), checksum = md5_out, row.names = NULL)
```

The checksum (MD5 hashes) of the exported files are:  
```{r, echo = FALSE}
info_out
```


---

# sessionInfo() and RStudio version

```{r}
sessionInfo()
```

RStudio version `r readLines("analysis/scripts/RStudioVersion.txt", n = 1)`.


---


END OF SCRIPT
