title: "Plot - Edge angle analysis - experimental data"
author: "Lisa Schunk"
date: "`r Sys.time()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: cerulean
    highlight: pygments
    number_sections: true
  github_document: 
    toc: true
    toc_depth: 3
    html_preview: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_format = "all", knit_root_dir = rprojroot::find_rstudio_root_file()) })

---

```{r Knitr Options, include = FALSE}
knitr::opts_chunk$set(comment = NA, message = FALSE, indent = "", error = TRUE)
```

# Goal of the script
This script plots all edge angle values. The three experiments will be treated separated.  



```{r}
dir_in <- "analysis/derived_data/"
dir_out <- "analysis/plots"
```

Raw data must be located in `r paste0("~/", dir_in)`.  
Formatted data will be saved in `r paste0("~/", dir_out)`.

The knit directory for this script is the project directory.
---


# Load packages
```{r}
library(openxlsx)
library(readxl)
library(R.utils)
library(ggplot2)
library(tools)
library(tidyverse)
library(patchwork)
library(ggsci)
library(ggfortify)
library(wesanderson)
library(doBy)
library(ggfortify)

```


---

# Get name, path and information of the file 
```{r}
data_file <- list.files(dir_in, pattern = "\\.xlsx$", full.names = TRUE)
md5_in <- md5sum(data_file)
info_in <- data.frame(file = basename(names(md5_in)), checksum = md5_in, row.names = NULL)
```


The checksum (MD5 hashes) of the imported file is:  
```{r, echo = FALSE}
info_in
```


# Load data into R object
```{r}
imp_data <- read.xlsx(xlsxFile = data_file, sheet = 1, startRow = 1, colNames = TRUE,
                      rowNames = FALSE, skipEmptyRows = FALSE) 
str(imp_data)
```

The imported file is: "`r paste0("~/", data_file)`"  

---


# Prepare variables
## Define numeric variables
```{r}
#num.var <- 12
num.var <- 14:length(imp_data)
```

The following variables will be used: 

```{r, echo=FALSE}
for (i in num.var) cat("[",i,"] ", names(imp_data)[i], "\n", sep="")
```


---


# Facet wrap to plot each experiment individually 
## Artfifical VS natural experiment 
```{r}
# selects only the data from the aVSn experiment   
aVSn <- filter(imp_data, experiment == "'aVSn'-experiment")
aVSn <- filter(aVSn, Angle_number == "3" | Angle_number == "4" | Angle_number == "5" | Angle_number == "6" | Angle_number == "7" | Angle_number == "8")

# adds a column that combines sample and location
aVSn_data <- unite(aVSn, ID_cycle, c(ID, strokes), remove = FALSE)

# computes the mean per sample 
aVSn_mean <- summaryBy(. ~ ID_cycle+ID+strokes+Contact.material+Raw.material+Edge.angle, data = aVSn_data, FUN = mean)

# gets new order 
aVSn_mean$strokes <- factor(aVSn_mean$strokes, levels=c("before", "50strokes", "250strokes", "1000strokes", "2000strokes"))

# gets the min/max range of the data set 
range_var <- range(aVSn_mean[["Three_point.mean"]]) 

# plots 
# plots first the lydit samples
p_lydite <- ggplot(data = aVSn_mean[grep("LYDIT", aVSn_mean[["ID_cycle"]]), ], aes(x = strokes, y = Three_point.mean, colour = Edge.angle)) + 
            geom_point(size = 2) +
            geom_line(aes(group = ID)) +
            facet_wrap(ID ~ Contact.material, nrow = 2)+
            theme_classic()+
            xlab("strokes") + ylab(NULL) +
            labs(colour = "Edge angle") +
            coord_cartesian(ylim = range_var) + 
            scale_colour_manual(values = "#CB2314") +
            scale_x_discrete(breaks = c("before", "50strokes", "250strokes" , "1000strokes", "2000strokes"), labels = c("0", "50", "250", "1000", "2000"))


# plots the flint samples 
p_flint <- ggplot(data = aVSn_mean[grep("FLT", aVSn_mean[["ID_cycle"]]), ], aes(x = strokes, y = Three_point.mean, colour = Edge.angle)) + 
            geom_point(size = 2) +
            geom_line(aes(group = ID)) +
            facet_wrap(ID ~ Contact.material, nrow = 2)+
            theme_classic()+
            xlab("strokes") + ylab(NULL) +
            labs(colour = "Edge angle") +
            coord_cartesian(ylim = range_var) +
            scale_colour_manual(values = "#CB2314") +
            scale_x_discrete(breaks = c("before", "50strokes", "250strokes" , "1000strokes", "2000strokes"), labels = c("0", "50", "250", "1000", "2000"))
print(p_flint)
	
# combines the flint and the lydite plots 
p <- p_flint + p_lydite + plot_layout(guides = 'collect')  
       
print(p)

#save to PDF
	file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_aVSn_EA_plot", ".pdf")
	ggsave(filename = file_out, plot = p, path = dir_out, device = "pdf",width = 250, height = 170, units = "mm")


```


## Tool function experiment 
```{r}
# selects only the data from the TFE experiment and the 'Three-point' method  
#TFE <- imp_data[7460:20629,1:14]
TFE <- filter(imp_data, experiment == "tool_function-experiment_cutting" | experiment == "tool_function-experiment_carving")
TFE <- filter(TFE, Angle_number == "3" | Angle_number == "4" | Angle_number == "5" | Angle_number == "6" | Angle_number == "7" | Angle_number == "8")

# adds a column that combines sample and location
TFE_data <- unite(TFE, ID_cycle, c(ID, strokes), remove = FALSE)
TFE_mean <- summaryBy(. ~ ID_cycle+ID+strokes+Contact.material+Raw.material+Edge.angle+Task, data = TFE_data, FUN = mean)

# removes the before from the sample FLT8-7 because these measurements are wrong -> this line needs to be removed when the correct data is there 
# computes the mean per sample 
TFE_final <- TFE_mean[c(0:49, 51:80, 83:122),]
# removes the test sample LYDIT5-14 (carving) 

# gets new order 
TFE_final$strokes <- factor(TFE_final$strokes, levels=c("before", "50strokes", "250strokes", "1000strokes", "2000strokes"))

# plots 
# plots first the lydite samples
p_lydite2 <- ggplot(data = TFE_final[grep("LYDIT", TFE_final[["ID_cycle"]]), ], aes(x = strokes, y = Three_point.mean, colour = Edge.angle)) + 
             geom_point(size = 2) +
             geom_line(aes(group = ID)) +
             facet_wrap(ID ~ Contact.material+Task, nrow = 3)+
             theme_classic()+
             xlab("strokes") + ylab(NULL) +
             labs(colour = "Edge angle") +
             ylim(30, 80) + 
             scale_colour_manual(values = c("#046C9A", "#FAD510")) +
             scale_x_discrete(breaks = c("before", "50strokes", "250strokes" , "1000strokes", "2000strokes"), labels = c("0", "50", "250", "1000", "2000"))


# plots the flint samples 
p_flint2 <- ggplot(data = TFE_final[grep("FLT", TFE_final[["ID_cycle"]]), ], aes(x = strokes, y = Three_point.mean, colour = Edge.angle)) + 
            geom_point(size = 2) +
            geom_line(aes(group = ID)) +
            facet_wrap(ID ~ Contact.material+Task, nrow = 3)+
            theme_classic()+
            xlab("strokes") + ylab(NULL) +
            labs(colour = "Edge angle") +
            ylim(30, 80) + 
            #coord_cartesian(ylim = range_var) +
            scale_colour_manual(values = c("#046C9A", "#FAD510")) +
            scale_x_discrete(breaks = c("before", "50strokes", "250strokes" , "1000strokes", "2000strokes"), labels = c("0", "50", "250", "1000", "2000"))

	
# combines the flint and the lydite plots 
p2 <- p_flint2 + p_lydite2 + plot_layout(guides = 'collect')  
       
print(p2)

#save to PDF
	file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_TFE_EA_plot", ".pdf")
	ggsave(filename = file_out, plot = p2, path = dir_out, device = "pdf", width = 370, height = 280, units = "mm")


```


## Initial experiment 
```{r}
# selects only the data from the initial experiment  
IE <- filter(imp_data, experiment == "initial_experiment")
# filters the first two and the two last sections out 
IE <- filter(IE, Angle_number == "3" | Angle_number == "4" | Angle_number == "5" | Angle_number == "6" | Angle_number == "7" | Angle_number == "8")

# adds a column that combines sample and location
IE_data <- unite(IE, ID_cycle, c(ID, strokes), remove = FALSE)

# computes the mean per sample 
IE_mean <- summaryBy(. ~ ID_cycle+ID+strokes+Contact.material+Raw.material+Edge.angle+Task, data = IE_data, FUN = mean)

# gets new order 
IE_mean$strokes <- factor(IE_mean$strokes, levels=c("before", "50strokes", "250strokes", "1000strokes", "2000strokes"))


# plots 
# plots first the lydit samples
p_lydite3 <- ggplot(data = IE_mean[grep("LYDIT", IE_mean[["ID_cycle"]]), ], aes(x = strokes, y = Three_point.mean, colour = Edge.angle)) + 
            geom_point(size = 2) +
            geom_line(aes(group = ID)) +
            facet_wrap(ID ~ Contact.material+Task, nrow = 3)+
            theme_classic()+
            xlab("strokes") + ylab(NULL) +
            labs(colour = "Edge angle") +
            ylim(35, 70) + 
            scale_colour_manual(values = c("#354823", "#CB2314")) +
            scale_x_discrete(breaks = c("before", "50strokes", "250strokes" , "1000strokes", "2000strokes"), labels = c("0", "50", "250", "1000", "2000"))
#print(p_lydite3)

# plots the flint samples 
p_flint3 <- ggplot(data = IE_mean[grep("FLT", IE_mean[["ID_cycle"]]), ], aes(x = strokes, y = Three_point.mean, colour = Edge.angle)) + 
            geom_point(size = 2) +
            geom_line(aes(group = ID)) +
            facet_wrap(ID ~ Contact.material+Task, nrow = 3)+
            theme_classic()+
            xlab("strokes") + ylab(NULL) +
            labs(colour = "Edge angle") +
            ylim(35, 70) + 
            scale_colour_manual(values = c("#354823", "#CB2314")) +
            scale_x_discrete(breaks = c("before", "50strokes", "250strokes" , "1000strokes", "2000strokes"), labels = c("0", "50", "250", "1000", "2000"))


# combines the flint and the lydite plots 
p3 <- p_flint3 + p_lydite3 + plot_layout(guides = 'collect')
       
print(p3)

# temporary solution
#save to PDF
	file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_IE1_EA_plot", ".pdf")
	ggsave(filename = file_out, plot = p_lydite3, path = dir_out, device = "pdf",width = 250, height = 170, units = "mm")

  file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_IE2_EA_plot", ".pdf")
	ggsave(filename = file_out, plot = p_flint3, path = dir_out, device = "pdf",width = 250, height = 170, units = "mm")
	
	
```


---


## Show files information
```{r}
files_out <- c(paste0(dir_out, file_out, ".xlsx"), paste0(dir_out, file_out, ".Rbin"))
md5_out <- md5sum(files_out)
info_out <- data.frame(files = basename(names(md5_out)), checksum = md5_out, 
                       row.names = NULL)
```


---

The checksum (MD5 hashes) of the exported files are:  
```{r, echo = FALSE}
info_out
```


---

# sessionInfo() and RStudio version

```{r}
sessionInfo()
```

RStudio version `r readLines("analysis/scripts/RStudioVersion.txt", n = 1)`.


---

END OF SCRIPT
